Physiological plasticity and variance in physiological rates vary across ectotherm populations

### 2024-07-13 ms.qmd
Updating code and figures to move lnVR in the main MS and lnCVR in supplement as it's jut too difficult to interpret lnCVR and we are alerady analysing lnRR (mean) so better to keep the variance lnVR bedide it so that it's cleaer and connects better to readers. CVR is roughly the same anyway

---
title: "Limited physiological plasticity and reduced relative variance in physiological rates of ectotherm populations under climate change"
bibliography: ./bib/refs.bib
csl: ./bib/ecology-letters.csl
format:
  docx:
    reference-doc: ./bib/template.docx
    output-file: "ms_revised.docx"
editor_options: 
  chunk_output_type: console
execute:
  freeze: auto  # re-render only when source changes
  cache: false
  echo: false
  warning: false
  error: false
  include: true
crossref:
  fig-title: 'Figure'
  fig-labels: arabic
  title-delim: "-"
  fig-prefix: "Figure"
  tbl-prefix: "Table"
editor: 
  markdown: 
    wrap: 72
---

@@

@eq-lnq10VR and @eq-slnq10VR describe the change in physiological rate
variance (@eq-lnq10VR) across a 10°C temperature change along with its
sampling variance (@eq-slnq10VR). While this is a useful metric, as
discussed by @Nakagawa2015-fo there is often a strong mean-variance
relationship that needs to be accounted for in analysing changes in
variance. As such, we calculated the coefficient of variation, which
standardizes changes in variance for changes in means as follows:

@@

where $CV$ is the coefficient of variation defined as $SD / R$. We refer
to $lnCVR_{Q_{10}}$ as relative variance because variance changes are
relative to the mean.

@@

$R_{1}$, $R_{2}$, $SD_{1}^2$ and $SD_{2}^2$ can all be taken from
samples of organisms measured acutely at two temperatures or after
having been acclimated these same temperatures ([@fig-concept]A). For
studies that measure acute and acclimated responses we used the mean,
standard deviation, and sample size to derive both acute and acclimation
$lnRR_{Q_{10}}$, $lnVR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ estimates. For
studies that only measured $R_{1}$, $R_{2}$, $SD_{1}^2$ and $SD_{2}^2$
acutely we could only calculate acute versions of these effect size
estimates. For all effect sizes the higher temperature was in the
numerator and the lower of the two temperatures in the denominator. As
such, positive effect sizes suggest that the mean (i.e.,
$lnRR_{Q_{10}}$) or relative variance (i.e., $lnCVR_{Q_{10}}$) is larger
at the higher of the two temperatures (numerator) when standardized to
10°C. Importantly, $lnRR_{Q_{10}}$, $lnCVR_{Q_{10}}$ and
$lnVR_{Q_{10}}$, as with $Q_{10}$ more generally, all assume that the
effect of temperature on physiological rates (or changes in relative
variance) is log-linear (see [@fig-concept]B & *Supplementary Materials* for further
discussion). We test and control for any violations of these assumptions
in our analysis (see below).

@@

@@

        # Contrast statistics
            contrast_lnrr_trait_stats <- data.frame( habitat = rep(c("t", "m", "f"), each = 2),
                                                     trait = rep(c("tissue/organism", "biochemical"), each = 1),
                                                    diff = sapply(data.frame(contrast_lnrr_trait_post), mean),
                                                     l95 = sapply(data.frame(contrast_lnrr_trait_post), function(x) quantile(x, 0.025)),
                                                     u95 = sapply(data.frame(contrast_lnrr_trait_post), function(x) quantile(x, 0.975)),
                                                       p = sapply(data.frame(contrast_lnrr_trait_post), function(x) p_value(pmcmc(x))))

###########################################################
## Model 1.4 - whether chance between q10 depends on life stage and habitat
###########################################################
rerun = FALSE
if(rerun){
  # More realistic model with 
  model1.4 <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type*habitat*life.hist, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor), dfs = "contain", 
                              test = "t", data = lnRRQ10_data)

@@

@@

   # Calculate the design matrix using this new data. Need to be careful here to check that the levels are set up the same as the posterior.

@@

```{r, fig-2}
#| label: fig-2
#| fig-cap: "A) Mean acute and acclimation $lnRR_{Q_{10}}$ across ectotherms in marine, freshwater, and terrestrial habitats. B) Mean acute and acclimation $lnCVR_{Q_{10}}$ across ectotherms in marine, freshwater and terrestrial habitats. For both plots, k = total number of effect size estimates while the numbers in brackets indicate the number of species. Sample sizes are the same for panel A) and B). In panels A) and B), thick bars are 95% confidence intervals (CI) and thin bars 95% prediction intervals (PI). $\\Beta$ values are the contrasts between acute and acclimation means within each habitat. $\\mu$ values are the overall meta-analytic means averaged across acute and acclimation types within each habitat type. In both cases, their 95% CI's are indicated within square brackets. $p_{MCMC}$ values are the posterior probability of the contrast or overall meta-analytic mean being different from zero. For ease of visualisation, all the raw data plotted for both acute and acclimation type effect sizes are presented as circles."
#| warning: false
#| echo: false

@@

##################
  # Figure 3b plot
################## 
  # Clear from the model that there really is no difference between the change in variance for acute or acclimation CVR. But, there is a clear reduction in between individual variance for terrestrial species. As temperature get hotter then the variance in physiological rates decreases. This is not true of aquatic systems.

  k_cvr <- manual_extract(model2.2, mod = "habitat", group = "record_num", exclude = NULL)

@@

### *Relative variation in physiological rates is reduced in terrestrial and marine ectotherms*

Relative variance in physiological rates ($lnCVR_{Q_{10}}$) showed a
decrease with increasing temperature across all habitat types, which was
especially pronounced in terrestrial and marine ectotherms. The total
proportion of heterogeneity in $lnCVR_{Q_{10}}$ was lower compared to
$lnRR_{Q_{10}}$ ($I^2_{sv}$ = `r mean(tot_i2_cvr)`, 95% CI:
`r quantile(tot_i2_cvr, 0.025)` to `r quantile(tot_i2_cvr, 0.975)`),
with most variation being driven by between-study and trait differences
("Model 1"; see Supplement @fig-s3). There were no differences between
$lnCVR_{Q_{10_{acute}}}$ and $lnCVR_{Q_{10_{acclimation}}}$. Reductions
in variance were driven by slower increases in the variance in
physiological rates as mean physiological rates increased with scaling
relationships between log standard deviation and log mean all being
smaller than 1 and varying across habitats (see *Supplementary
Materials*; @tbl-slopes & @fig-mv). There was a `r mean(t_cvr_mean_ef)`%
(95% CI: `r quantile(t_cvr_mean_ef, 0.025)[1]` to
`r quantile(t_cvr_mean_ef, 0.975)[1]`, $p_{MCMC}$ =
`r pmcmc(t_cvr_mean)`) reduction in relative physiological rate variance
for terrestrial ectotherms and a `r mean(m_cvr_mean_ef)`% (95% CI:
`r quantile(m_cvr_mean_ef, 0.025)[1]` to
`r quantile(m_cvr_mean_ef, 0.975)[1]`, $p_{MCMC}$ =
`r pmcmc(m_cvr_mean)`) reduction in relative variation for marine
ectotherms when temperatures increased by 10$^\circ$C. In contrast,
freshwater ectotherms exhibited a smaller reduction in relative
physiological rate variance at high temperatures
(`r mean(f_cvr_mean_ef)`%, 95% CI: `r quantile(f_cvr_mean_ef, 0.025)[1]`
to `r quantile(f_cvr_mean_ef, 0.975)[1]`, $p_{MCMC}$ =
`r pmcmc(f_cvr_mean)`) (results from "Model 2").

Across habitats whole-organism versus biochemical processes showed a
similar decrease in relative variance in physiological rates
([@fig-btrait]D-F; see *Supplementary Materials*, @fig-cvrtrait) for more detailed trait
analysis), except for biochemical processes in ectotherms occupying
freshwater habitats, which showed small, albeit significant increase in
relative variance when acclimated ([@fig-btrait]F). Reduced relative
variance was particularly prominent for resting metabolic rates and
sprint speed although traits differed in whether they exhibited a
reduction in variation in physiological rates at high temperatures (see
*Supplemental Material*, @fig-cvrtrait) ("Model 3"). Each life-history stage
exhibited the same pattern of relative variance change in each of the
habitats ("Model 4"; @fig-cvrlh).

```{r, fig-btrait}
#| label: fig-btrait
#| fig-cap: "Estimated mean acclimation and acute $lnRR_{Q_{10}}$ (A-C) and $lnCVR_{Q_{10}}$ (D-F) effect sizes for tissue/whole-orgamism traits and biochemical traits across marine (A & D), freshwater (B & E) and terrestrial (C & F) habitats. Across all plots, thick bars indicate 95% confidence intervals and thin bars indicate 95% prediction intervals. k = total number of effect size estimates while the numbers in brackets indicate the number of species. For ease of visualisation, raw data for both trait categories are presented but points are not distinguished by different symbols. "
#| echo: false
#| warning: false

@@

n_lnrr_btrait  <- lnRRQ10_data  %>% filter(trait_broad != "other") %>% 
              group_by(habitat, trait_broad, type) %>% summarise(k = n(), spp = length(unique(spp))) %>% arrange(trait_broad, type, habitat)
n_lncvr_btrait <- lnCVRQ10_data %>% filter(trait_broad != "other") %>% 
              group_by(habitat, trait_broad, type) %>% summarise(k = n(), spp = length(unique(spp))) %>% arrange(trait_broad, type, habitat)

@@

############
# Full fig
############
figbtrait <-(p3 | p1 | p2) / (p3_cvr | p1_cvr | p2_cvr) + plot_annotation(tag_levels = "A") 
ggsave(figbtrait, filename = "./output/figures/fig4.png", width = 28, height = 14, device = "png")

@@

```{r, fig-cvrlh}
#| label: fig-cvrlh
#| fig-cap: "Estimated mean acclimation and acute $lnRR_{Q_{10}}$ (A-C) and $lnCVR_{Q_{10}}$ (D-F) for adult (a) and juvenile (j) life-history stages for marine (A & D), freshwater (B & E) and terrestrial (C & F) ectotherms. Across all plots, thick bars indicate 95% confidence intervals and thin bars indicate 95% prediction intervals. k = total number of effect size estimates while the numbers in brackets indicate the number of species. For ease of visualisation, raw data for both adult and juvenile life-history stages are presented but points are not distinguished by different symbols. $\\Beta$ values are the contrasts between acute and acclimation means within each life stage. $p_{MCMC}$ values are the posterior probability of the contrast being different from zero."
#| echo: false
#| warning: false

@@

n_lnrr_lh  <- lnRRQ10_data  %>% group_by(habitat, life.hist, type) %>% summarise(k = n(), spp = length(unique(spp))) %>% arrange(type, life.hist, habitat)
n_lncvr_lh <- lnCVRQ10_data %>% group_by(habitat, life.hist, type) %>% summarise(k = n(), spp = length(unique(spp))) %>% arrange(type, life.hist, habitat)

@@

fig4 <-(p3 | p1 | p2) / (p3_cvr | p1_cvr | p2_cvr) + plot_annotation(tag_levels = "A") 
ggsave(fig4, filename = "./output/figures/fig5.pdf", width = 28, height = 14, device = "pdf")

@@

@@

@@

@@

#### *Acute and acclimation* $lnRR_{Q_{10}}$ for detailed trait categories across marine, freshwater and terrestrial taxa

@@

#### *Comparing raw variance changes using* $lnVR_{Q_{10}}$

@@

        # Contrast statistics
            contrast_lnvr_trait_stats <- data.frame( habitat = rep(c("t", "m", "f"), each = 2),
                                                     trait = rep(c("tissue/organism", "biochemical"), each = 1),
                                                    diff = sapply(data.frame(contrast_lnvr_trait_post), mean),
                                                     l95 = sapply(data.frame(contrast_lnvr_trait_post), function(x) quantile(x, 0.025)),
                                                     u95 = sapply(data.frame(contrast_lnvr_trait_post), function(x) quantile(x, 0.975)),
                                                       p = sapply(data.frame(contrast_lnvr_trait_post), function(x) p_value(pmcmc(x))))

>>>>

---
title: "Physiological plasticity and variance in physiological rates vary across ectotherm populations"
bibliography: ./bib/refs.bib
csl: ./bib/ecology-letters.csl
format:
  docx:
    reference-doc: ./bib/template.docx
    output-file: "ms_revised.docx"
editor_options: 
  chunk_output_type: console
execute:
  freeze: auto  # re-render only when source changes
  cache: false
  echo: false
  warning: false
  error: false
  include: true
crossref:
  fig-title: 'Figure'
  fig-labels: arabic
  title-delim: "-"
  fig-prefix: "Figure"
  tbl-prefix: "Table"
editor: 
  markdown: 
    wrap: 72
---

@@

where parameters are defined above. @eq-lnq10VR and @eq-slnq10VR describe the change in physiological rate
variance (@eq-lnq10VR) across a 10°C temperature change along with its
sampling variance (@eq-slnq10VR). As
discussed by @Nakagawa2015-fo there is often a strong mean-variance
relationship. As such, the coefficient of variation is often used because it allows one to
standardize changes in variance as the mean changes as follows:

@@

where $CV$ is the coefficient of variation defined as $SD / R$. We refer
to $lnCVR_{Q_{10}}$ as relative variance because variance changes are
relative to the mean. While we analyse $lnCVR_{Q_{10}}$ it does make certain assumptions about mean-variance relationships, and given that we are analysing the mean alongside the variance we present results on $lnCVR_{Q_{10}}$ in the supplement.

@@

$R_{1}$, $R_{2}$, $SD_{1}^2$ and $SD_{2}^2$ can all be taken from
samples of organisms measured acutely at two temperatures or after
having been acclimated these same temperatures ([@fig-concept]A). For
studies that measure acute and acclimated responses we used the mean,
standard deviation, and sample size to derive both acute and acclimation
$lnRR_{Q_{10}}$, $lnVR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ estimates. For
studies that only measured $R_{1}$, $R_{2}$, $SD_{1}^2$ and $SD_{2}^2$
acutely we could only calculate acute versions of these effect size
estimates. For all effect sizes the higher temperature was in the
numerator and the lower of the two temperatures in the denominator. As
such, positive effect sizes suggest that the mean (i.e.,
$lnRR_{Q_{10}}$), variance ($lnVR_{Q_{10}}$) or relative variance (i.e., $lnCVR_{Q_{10}}$) is larger
at the higher of the two temperatures (numerator) when standardized to
10°C. Importantly, our effect sizes, as with $Q_{10}$ more generally, all assume that the
effect of temperature on physiological rates (or changes in
variance) is log-linear (see [@fig-concept]B & *Supplementary Materials* for further
discussion). We test and control for any violations of these assumptions
in our analysis (see below).

@@

  # Calculated weighted mean by sample size. 
    f_RR_mean <- (f_acute * N[1,"n"][[1]] + f_acclim * N[2,"n"][[1]]) / sum(N[1:2, "n"]); pmcmc(f_RR_mean)
    m_RR_mean <- (m_acute * N[3,"n"][[1]] + m_acclim * N[3,"n"][[1]]) / sum(N[3:4, "n"]); pmcmc(m_RR_mean)
    t_RR_mean <- (t_acute * N[5,"n"][[1]] + t_acclim * N[6,"n"][[1]]) / sum(N[5:6, "n"]); pmcmc(t_RR_mean)

@@

        # Contrast statistics
            contrast_lnrr_trait_stats <- data.frame( habitat = rep(c("t", "m", "f"), each = 2),
                                                     trait = rep(c("tissue/organism", "biochemical"), each = 1),
                                                    diff = sapply(data.frame(contrast_lnrr_trait_post), mean),
                                                     l95 = sapply(data.frame(contrast_lnrr_trait_post), function(x) quantile(x, 0.025)),
                                                     u95 = sapply(data.frame(contrast_lnrr_trait_post), function(x) quantile(x, 0.975)),
                                                       p = sapply(data.frame(contrast_lnrr_trait_post), function(x) p_value(pmcmc(x))))
       # Contrasts for pooled acute and acclimation across trait categories

            contrast_lnrr_trait_cat_stats <- rbind(get_whole_bio(pred_lnrr_trait_post, habitat = "t", trait_b_data),
                                                        get_whole_bio(pred_lnrr_trait_post, habitat = "m", trait_b_data),
                                                        get_whole_bio(pred_lnrr_trait_post, habitat = "f", trait_b_data))
###########################################################
## Model 1.4 - whether chance between q10 depends on life stage and habitat
###########################################################
rerun = FALSE
if(rerun){
  # More realistic model with 
  model1.4 <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type*habitat*life.hist, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor), dfs = "contain", 
                              test = "t", data = lnRRQ10_data)

@@

  # Now, lets get contrasts between habitats averaged over type
      f_m_c <- f_VR_mean - m_VR_mean   ; pmcmc(f_m_c); mean(f_m_c)
      f_t_c <- t_VR_mean_ef - f_VR_mean; pmcmc(f_t_c); mean(f_t_c)
      m_t_c <- t_VR_mean_ef - m_VR_mean; pmcmc(m_t_c); mean(m_t_c)

@@

        # Contrast statistics
            contrast_lnvr_trait_stats <- data.frame( habitat = rep(c("t", "m", "f"), each = 2),
                                                     trait = rep(c("tissue/organism", "biochemical"), each = 1),
                                                    diff = sapply(data.frame(contrast_lnvr_trait_post), mean),
                                                     l95 = sapply(data.frame(contrast_lnvr_trait_post), function(x) quantile(x, 0.025)),
                                                     u95 = sapply(data.frame(contrast_lnvr_trait_post), function(x) quantile(x, 0.975)),
                                                       p = sapply(data.frame(contrast_lnvr_trait_post), function(x) p_value(pmcmc(x))))
        # Contrasts for pooled acute and acclimation across trait categories

            contrast_lnvr_trait_cat_stats <- rbind(get_whole_bio(pred_lnvr_traitb_post, habitat = "t", trait_b_data),
                                                        get_whole_bio(pred_lnvr_traitb_post, habitat = "m", trait_b_data),
                                                        get_whole_bio(pred_lnvr_traitb_post, habitat = "f", trait_b_data))

@@

```{r, fig-2}
#| label: fig-2
#| fig-cap: "A) Mean acute and acclimation $lnRR_{Q_{10}}$ across ectotherms in marine, freshwater, and terrestrial habitats. B) Mean acute and acclimation $lnVR_{Q_{10}}$ across ectotherms in marine, freshwater and terrestrial habitats. For both plots, k = total number of effect size estimates while the numbers in brackets indicate the number of species. Sample sizes are the same for panel A) and B). In panels A) and B), thick bars are 95% confidence intervals (CI) and thin bars 95% prediction intervals (PI). $\\Beta$ values are the contrasts between acute and acclimation means within each habitat. $\\mu$ values are the overall meta-analytic means averaged across acute and acclimation types within each habitat type. In both cases, their 95% CI's are indicated within square brackets. $p_{MCMC}$ values are the posterior probability of the contrast or overall meta-analytic mean being different from zero. For ease of visualisation, all the raw data plotted for both acute and acclimation type effect sizes are presented as circles."
#| warning: false
#| echo: false

@@

##################
  # Figure 3b plot
################## 
    # Sensitivity analysis that looks at how variance changes but not controlling for the mean.
  k_vr <- manual_extract(model3.2, mod = "habitat", group = "spp", exclude = NULL)

  p3b <- orchard_plot(model3.2, xlab = TeX("$lnVR_{Q_{10}}$"), mod = "habitat", by = "type", group  = "record_num", weights = "prop", angle = 45, trunk.size = 1.8, branch.size = 2.8) +
              scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) + ylim(-2,3) +

        annotate("text", x = 3.5, y = -2, label =TeX(paste0("$\\mu = $", round(mean(t_VR_mean), 2), " [", round(quantile(t_VR_mean, 0.025), 2)," to ", round(quantile(t_VR_mean, 0.975),2), "] ", " $\\,p_{MCMC}$ = ", p_value(round(pmcmc(t_VR_mean), digit = 2)))), hjust = 0, size = 8) + 
    annotate("text", x = 3.2, y = 1.5, label =TeX(paste0("$\\beta = $", "NS")), hjust = 0, size = 8) +
  annotate("text", x = 3.20, y = -2, label =paste0(round(mean(t_VR_mean_ef), 1), "%"), hjust = 0, size = 8) + labs(shape = "Type") + scale_fill_manual(values = c(col)) + annotate("text", y = 2.5, x = c(1.3, 2.3, 3.3)+0.1, label = TeX(paste0("\\it{k} = ", k_vr$k, " (", k_vr$sdy, ")")), size = 6) + scale_fill_manual(values = col) + scale_color_manual(values = rep("black",3))

  p3b <- p3b + theme_classic() + theme(axis.title  = element_text(size = 24),
                                       axis.text = element_text(size = 24),
                                       axis.text.y = element_text(angle = 45, size = 24),
                                       legend.title = element_text(size = 30),
                               legend.text = element_text(size = 24),
                               title = element_text(size = 24), legend.position = "top")

@@

### *Variation in physiological rates increases in aquatic but not terrestrial ectotherms*

Variance in physiological rates ($lnVR_{Q_{10}}$) showed an
increase with increasing temperature across all habitat types. However, physiological rate variance increase significantly more in freshwater (freshwater vs terrestrial: $\beta$ = `r mean(f_t_c)`, 95% CI: `r quantile(f_t_c, 0.025)` to `r quantile(f_t_c, 0.975)`, $p_{MCMC}$ = `r p_value(pmcmc(f_t_c))`) and marine ectotherms (marine vs terrestrial: $\beta$ = `r mean(m_t_c)`, 95% CI: `r quantile(m_t_c, 0.025)` to `r quantile(m_t_c, 0.975)`, $p_{MCMC}$ = `r p_value(pmcmc(m_t_c))`) compared to the limited increase in terrestrial ecotherms (@fig-2). The total
proportion of heterogeneity in $lnVR_{Q_{10}}$ was lower compared to
$lnRR_{Q_{10}}$ ($I^2_{sv}$ = `r mean(tot_i2_VR)`, 95% CI:
`r quantile(tot_i2_VR, 0.025)` to `r quantile(tot_i2_VR, 0.975)`),
with most variation being driven by between-study and trait differences
("Model 1"; see *Supplementary Materials* @fig-s3). There were no differences between
$lnVR_{Q_{10_{acute}}}$ and $lnVR_{Q_{10_{acclimation}}}$. Reductions
in variance were driven by slower increases in the variance in
physiological rates as mean physiological rates increased with scaling
relationships between log standard deviation and log mean all being
smaller than 1 and varying across habitats (see *Supplementary
Materials*; @tbl-slopes & @fig-mv). There was a `r mean(t_VR_mean_ef)`%
(95% CI: `r quantile(t_VR_mean_ef, 0.025)[1]` to
`r quantile(t_VR_mean_ef, 0.975)[1]`, $p_{MCMC}$ =
`r pmcmc(t_VR_mean)`) reduction in relative physiological rate variance
for terrestrial ectotherms and a `r mean(m_VR_mean_ef)`% (95% CI:
`r quantile(m_VR_mean_ef, 0.025)[1]` to
`r quantile(m_VR_mean_ef, 0.975)[1]`, $p_{MCMC}$ =
`r pmcmc(m_VR_mean)`) reduction in relative variation for marine
ectotherms when temperatures increased by 10$^\circ$C. In contrast,
freshwater ectotherms exhibited a smaller reduction in relative
physiological rate variance at high temperatures
(`r mean(f_VR_mean_ef)`%, 95% CI: `r quantile(f_VR_mean_ef, 0.025)[1]`
to `r quantile(f_VR_mean_ef, 0.975)[1]`, $p_{MCMC}$ =
`r pmcmc(f_VR_mean)`) (results from "Model 2"). The relative differences between habitats were similar when considering $lnCVR_{Q_{10}}$ (See *Supplementary Materials*).

Across habitats whole-organism versus biochemical processes showed a
similar change in variance in physiological rates
([@fig-btrait]D-F; see *Supplementary Materials*, @fig-cvrtrait) for more detailed trait
analysis), except for biochemical processes in ectotherms occupying
freshwater habitats, which showed small, albeit significant increase in
relative variance when acclimated ([@fig-btrait]F). Reduced relative
variance was particularly prominent for resting metabolic rates and
sprint speed although traits differed in whether they exhibited a
reduction in variation in physiological rates at high temperatures (see
*Supplemental Material*, @fig-cvrtrait) ("Model 3"). Each life-history stage
exhibited the same pattern of relative variance change in each of the
habitats ("Model 4"; @fig-cvrlh).

```{r, fig-btrait}
#| label: fig-btrait
#| fig-cap: "Estimated mean acclimation and acute $lnRR_{Q_{10}}$ (A-C) and $lnVR_{Q_{10}}$ (D-F) effect sizes for tissue/whole-orgamism traits and biochemical traits across marine (A & D), freshwater (B & E) and terrestrial (C & F) habitats. Across all plots, thick bars indicate 95% confidence intervals and thin bars indicate 95% prediction intervals. k = total number of effect size estimates while the numbers in brackets indicate the number of species. For ease of visualisation, raw data for both trait categories are presented but points are not distinguished by different symbols. "
#| echo: false
#| warning: false

@@

n_lnrr_btrait  <- lnRRQ10_data  %>% filter(trait_broad != "other") %>% 
              group_by(habitat, trait_broad, type) %>% summarise(k = n(), spp = length(unique(spp))) %>% arrange(trait_broad, type, habitat)
n_lnvr_btrait <- lnVRQ10_data %>% filter(trait_broad != "other") %>% 
              group_by(habitat, trait_broad, type) %>% summarise(k = n(), spp = length(unique(spp))) %>% arrange(trait_broad, type, habitat)

@@

############
# lnVR
############
m_r_vr <- n_lnvr_btrait %>% filter(habitat == "m")
f_r_vr <- n_lnvr_btrait %>% filter(habitat == "f")
t_r_vr <- n_lnvr_btrait %>% filter(habitat == "t")

t_trait_b_vr <- contrast_lnvr_trait_stats %>% filter(habitat == "t") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(trait)
m_trait_b_vr <- contrast_lnvr_trait_stats %>% filter(habitat == "m") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(trait)
f_trait_b_vr <- contrast_lnvr_trait_stats %>% filter(habitat == "f") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(trait)

p1_vr_trait <- orchard_plot(model3.3_m_broad, mod = "trait_broad", at = list(c_accl_time = 0, habitat = "m"), group = "record_num", xlab = TeX("$lnVR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, angle = 45, branch.size = branch.size, trunk.size = trunk.size) + labs(title = "Marine") + add_theme() + ylim(-1.5, 3) + theme(legend.position = "none", axis.text.y = element_blank()) + annotate("text", x = x.pos, y = 2.5, label = TeX(paste0("\\it{k} = ", m_r_vr$k, " (", m_r_vr$spp, ")")), size = size_annot) + scale_fill_manual(values = rep(col[2],4)) + scale_color_manual(values = rep("black",4)) + annotate("text", x = yloc, y = -1.5, label = TeX(paste0("$\\beta$ = ", m_trait_b_vr$diff, " [", m_trait_b_vr$l95, " to ", m_trait_b_vr$u95, "] ", "$\\,p_{MCMC}$ = ", m_trait_b_vr$p)), hjust = 0, size = 8)

p2_vr_trait <- orchard_plot(model3.3_f_broad, mod = "trait_broad", at = list(c_accl_time = 0, habitat = "f"), group = "record_num", xlab = TeX("$lnVR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size, angle = 45)+ labs(title = "Freshwater")  + add_theme() + theme(legend.position = "none", axis.text.y = element_blank()) + ylim(-1.5, 3) + annotate("text", x = x.pos, y = 2.5, label = TeX(paste0("\\it{k} = ", f_r_vr$k, " (", f_r_vr$spp, ")")), size = size_annot) + theme(legend.position = "right", legend.justification = c(1, 1)) + guides(shape = guide_legend(direction = "vertical"), size = guide_legend(direction = "vertical")) + scale_fill_manual(values = rep(col[1],4)) + scale_color_manual(values = rep("black", 4)) + annotate("text", x = yloc, y = -1.5, label = TeX(paste0("$\\beta$ = ", f_trait_b_vr$diff, " [", f_trait_b_vr$l95, " to ", f_trait_b_vr$u95, "] ", "$\\,p_{MCMC}$ = ", f_trait_b_vr$p)), hjust = 0, size = 8)

p3_vr_trait <- orchard_plot(model3.3_t_broad, mod = "trait_broad", at = list(c_accl_time = 0, habitat = "t", legend.pos = "none"), group = "record_num", xlab = TeX("$lnVR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size, angle = 45)+ labs(title = "Terrestrial") + add_theme()  +  ylim(-1.5, 3) + theme(legend.position = "none") + annotate("text", x = x.pos, y = 2.5, label = TeX(paste0("\\it{k} = ", t_r_vr$k, " (", t_r_vr$spp, ")")), size = size_annot) + scale_fill_manual(values = rep(col[3], 4)) + scale_color_manual(values = rep("black", 4)) + annotate("text", x = yloc, y = -1.5, label = TeX(paste0("$\\beta$ = ", t_trait_b_vr$diff, " [", t_trait_b_vr$l95, " to ", t_trait_b_vr$u95, "] ", "$\\,p_{MCMC}$ = ", t_trait_b_vr$p)), hjust = 0, size = 8)

############
# Full fig
############
figbtrait <-(p3 | p1 | p2) / (p3_vr_trait | p1_vr_trait | p2_vr_trait) + plot_annotation(tag_levels = "A") 
ggsave(figbtrait, filename = "./output/figures/fig4.png", width = 28, height = 14, device = "png")

@@

```{r, fig-cvrlh}
#| label: fig-cvrlh
#| fig-cap: "Estimated mean acclimation and acute $lnRR_{Q_{10}}$ (A-C) and $lnVR_{Q_{10}}$ (D-F) for adult (a) and juvenile (j) life-history stages for marine (A & D), freshwater (B & E) and terrestrial (C & F) ectotherms. Across all plots, thick bars indicate 95% confidence intervals and thin bars indicate 95% prediction intervals. k = total number of effect size estimates while the numbers in brackets indicate the number of species. For ease of visualisation, raw data for both adult and juvenile life-history stages are presented but points are not distinguished by different symbols. $\\Beta$ values are the contrasts between acute and acclimation means within each life stage. $p_{MCMC}$ values are the posterior probability of the contrast being different from zero."
#| echo: false
#| warning: false

@@

n_lnrr_lh  <- lnRRQ10_data  %>% group_by(habitat, life.hist, type) %>% summarise(k = n(), spp = length(unique(spp))) %>% arrange(type, life.hist, habitat)
n_lnvr_lh <- lnVRQ10_data %>% group_by(habitat, life.hist, type) %>% summarise(k = n(), spp = length(unique(spp))) %>% arrange(type, life.hist, habitat)

@@

fig4 <-(p3 | p1 | p2) / (p3_vr | p1_vr | p2_vr) + plot_annotation(tag_levels = "A") 
ggsave(fig4, filename = "./output/figures/fig5.pdf", width = 28, height = 14, device = "pdf")

@@

```{r wildlnvr}
#| label: wildlnvr
#| cache: true
#| include: false
#| echo: false
#| warning: false
#| 
#######################################
  ## lnCVRQ10 from Wild Populations ##
#######################################
##################################################
  # Organise the data and prune tree
##################################################

@@

# Subset the data
  lnVRQ10_data_wild <- data_long2 %>% 
    filter(grepl("lnVR", name) & !is.na(effect_size) & !is.na(V) & !V == 0 & !is.na(acclim_period) & source == "w") %>% 
                        mutate(spp = species_full, 
                                     c_accl_time = acclim_period - mean(acclim_period, na.rm = TRUE), 
                                     match = interaction(spp, lat, long),
                                     max_temp_c = scale(temp_2, scale = FALSE))

# Now, join climate data to full data. This will give us all the climate variability data for each site across time
  lnVRQ10_data_wild <- left_join(lnVRQ10_data_wild, climate2, by = "match")

@@

# Check and prune tree
  tree_lnvr_wild <- tree_checks(data = lnVRQ10_data_wild, tree = tree_p0.7, dataCol = "species_full", type = "prune")
  tree_checks(data = lnVRQ10_data_wild, tree = tree_lnvr_wild, dataCol = "species_full", type = "check") # Good

# Create the phylo correlation matrix
  phylo_cor_wild_vr <- vcv(tree_lnvr_wild, cor = TRUE)
          A_wild_vr <- vcv(tree_lnvr_wild, cor = FALSE) # For bayes, need cov matrix

# Scale environment variables
    lnVRQ10_data_wild <- lnVRQ10_data_wild %>% mutate(     cv_c = scale(cv),
                                                        acf_all_c = scale(acf_all),
                                                        acf_lag1_c = scale(acf_lag1))

####################################################
  # Model 5.3: Model environmental variability by acute/acclim
####################################################
rerun = FALSE
if(rerun){
  # More realistic model with 
  model5.3          <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + cv_c + type + type:cv_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_vr), dfs = "contain", 
                              test = "t", data = lnVRQ10_data_wild)

   model5.3_bayes <- brms::brm(effect_size | se(sqrt(V), sigma = TRUE) ~  max_temp_c + c_accl_time + cv_c + type + type:cv_c + (~1 | record_num) + (1 | gr(species_full, cov = A_wild_vr)) + (1 + max_temp_c| spp) + (1|trait), data2 = list(A_wild_vr=A_wild_vr), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, data = lnVRQ10_data_wild)

  # Save model5.3
  saveRDS(model5.3, file = "./output/models/lnVRQ10_5.3")
  saveRDS(model5.3_bayes, file = "./output/models/lnVRQ10_5.3_bayes")

  } else{
        model5.3 <- readRDS("./output/models/lnVRQ10_5.3")
  model5.3_bayes <- readRDS("./output/models/lnVRQ10_5.3_bayes")
}

@@

#### *Acute and acclimation* $lnRR_{Q_{10}}$ for detailed trait categories across marine, freshwater and terrestrial taxa (REMOVE)

@@

#### *Comparing relative variance changes using* $lnCVR_{Q_{10}}$

@@

   # Contrasts for pooled acute and acclimation across trait categories

            contrast_lncvr_trait_cat_stats <- rbind(get_whole_bio(pred_lncvr_traitb_post, habitat = "t", trait_b_data),
                                                        get_whole_bio(pred_lncvr_traitb_post, habitat = "m", trait_b_data),
                                                        get_whole_bio(pred_lncvr_traitb_post, habitat = "f", trait_b_data))
